Magalhaes and Schwilk Supplementary Materials

### 2019-08-06 ms/Magalhaes_Schwilk_litter_flam_moisture.tex
Correct SI fig and table numbers, add SI figures

% Fine fuel moisture content reduces ignition probability and flame spread rate \citep{Gisborne-1936,   Fons-1946, Anderson+Rothermal-1965}.

Litter packing (measured as litter-bed density or packing ratio) controls the
flammability of dry litter fuels in seasonally dry climates because higher
litter density reduces flammability. Leaf litter density is driven by leaf size
and shape \citep{Fonda+Belanger+etal-1998, Scarff+Westoby-2006,
  Kane+Varner+etal-2008}. These litter particle effects on litter bed aeration,
in turn, have effects on litter bed flammability
\citep{Ganteaume+Marielle+etal-2011, Schwilk+Caprio-2011,
  Kreye+Varner+etal-2013}. These species-specific effects are well established.
However, how moisture dynamics modulate species specific flammability is poorly
known, and could be driven by by leaf traits, emergent litter litter traits, or
both. Species specific effects on litter flammability are recorded for the
forests of the southeastern US \citep{Nowacki+Abrams-2008} where
species-specific litter drying rates play a role in flammability feedbacks
\citep{Kreye+Varner+etal-2013}. Similarly, in mixed-conifer forest of the
western USA, where the fire regime is characterized by surface fires fueled by
plant litter, overstory tree composition drives variation in local fire
behavior \citep{Schwilk+Caprio-2011} as a result of differential litter
flammability across tree species and as a result of synergistic interactions in
multi-species litter mixtures \citep{Magalhaes+Schwilk-2012}. These
flammability feedbacks may help explain historical fire frequency patterns, but
current estimates of litter flammability in mixed conifer forests are based
largely on burning dry litter.

@@

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (36 36N, 118 42 W) between 1600 and 2400 m elevation in
mixed-conifer forest. We chose eight dominant tree species that are
representative of this type of forest: \emph{Pinus jeffreyi} Grev. \& Balf.,
\emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider litter the top layer of leaves and small
twigs less than 0.625 cm diameter (= 1-hour fuel) that have fallen in the
previous year (mostly undecomposed). Leaf litter was collected in the summer
(mid-June to mid-July) in 2012 from 21 separate sites across the
study area. We collected from a minimum of four populations per species with
each population 9--32\,km distant from its neighbors. An exception to this was
\emph{A. magnifica} which is restricted to the higher elevations and could only
be collected from one area, due to designated wilderness boundaries and
logistical constraints. Within that area, \emph{A. magnifica} litter was
collected from over 10 different individuals. For the remaining species, the
collection involved 2--4 individual trees at each site, separated from one
another by at least 10\,m. The litter was collected 2\,m away from the trunk to
obtain a more uniform sample of leaf and fine branchlet litter, because bark
and heavier branches tend to fall closer to the tree trunk. All litter was air
dried to $< 5$\% fuel moisture by dry weight for weighing and dividing into
moisture trials. The moisture manipulations and subsequent burning trials were
conducted in April--August 2015.

@@

Where $m$ is moisture content at time $t$, and $\lambda$ is the exponential
decay coefficient. To estimate these coefficients we fit a linear mixed-effects
model by first taking the natural logarithm of moisture content, with time
since wetting and species as fixed effects nested within replicate basket. We
extracted the coefficients of the fitted model to describe maximum water
retainability ($m_{max}$; g of water per g of dry mass) and desiccation rate
($\lambda$; hour$^{-1}$) for each species. All analyses were conducted in R
\citep{RCoreTeam-2019}. We fit models with R packages `lme4'
\citep{Bates_Machler_etal-2015} and `afex' \citep{Singmann_Bolker_etal-2017}.
Approximate degrees of freedom and p-values were calculated by the
Kenward-Roger approximation \citep{Kenward_Roger-1997} which is recommended by
Luke \citeyear{Luke-2017} as producing acceptable Type 1 error rates. We tested if
one leaf trait (specific leaf area) and one litter trait (litter bulk density)
influenced moisture dynamics by fitting linear models with species means as
observations. For each response variable we fit a single linear model with two
predictors and an interaction term.

Our past work \citep{Magalhaes+Schwilk-2012} demonstrated that dry litter
flammability across these eight species was driven by leaf traits and litter
packing. These previously reported flammability trials demonstrated that
species within a genus tended to behave similarly and the two scale-leaved
genera in the Cupressaceae, \emph{Calocedrus} and \emph{Sequoiadendron}, also
behaved similarly to one another. Therefore, to preserve power we used taxonomic
groups (\emph{Abies}, Cupressaceae, \emph{Quercus}, and \emph{Pinus}) as our
flammability functional groups and treated these as fixed effects in our
models. The effect of fuel moisture and time since wetting on flammability
parameters were modeled using linear models with fuel moisture or time since
wetting and taxon as fixed effects and average vapor pressure deficit in the
burning room during that trial as a nuisance covariate. We used linear mixed
models to predict spread rate, but fuel consumption tends to exhibit threshold
effects with few intermediate values. Therefore, we used mixed generalized
linear models with a beta regression link with the `glmmADMB` package
\citep{Skaug_Fournier_etal-2016} to model percent fuel consumed as a function of
moisture content and of time since wetting.

We investigated possible non-additive effects on moisture and on flammability
by testing if moisture content and flammability of mixtures were predicted by
the average of the constituent species. To choose mixtures, we first grouped
the species according to their maximum water retention and drying rate
according to estimated marginal means by species using the R package `emmeans'
\citep{Lenth-2019}. This resulted in four groups of species and we then selected
one representative species from each litter type group and produced all four
possible three-species litter mixtures using those four species. For each
mixture and time since wetting, we predicted moisture content based on the mean
content of the three constituent species at that time since wetting. We
predicted flammability parameters based on the average flammability measures
from individual species trials at each time since wetting and then calculated a
predicted spread rate and fuel consumption as the average of these three values
for the mixture. The expectation from the null model was that the difference
between observed and predicted values of flammability will be zero. We tested
if there were significant departures from the null (zero) with a mixed effects
linear model. For flammability parameters, we fit linear mixed effect models
with the observed-predicted flammability value as the dependent variable and
used Wald tests to test if the centered intercept was different than zero (a
measure of overall non-additivity) and if the slope with time since wetting was
different from zero (indicating non-additivity changing with drying).

@@

There were marked differences in the extent of moisture species retain
initially following wetting and draining by gravity (all $P$ < 0.0001, Fig.
\ref{fig:drydown} and Table \ref{tab:drydown}). \emph{Q. kelloggii}, the only
broadleaf, is the species that retained the most moisture initially.
\emph{Quercus}, like the pines, had a rapid drying rate (Fig. \ref{fig:drydown}
and SI Fig. 1). Calculating estimated marginal means following the linear
mixed-effects model discriminated the species into three different maximum
water retention groups: 1) \emph{Q. kelloggii}; 2) \emph{P. lambertiana} and
\emph{C. decurrens}; and 3) \emph{A. magnifica}, \emph{S. giganteum}, \emph{A.
  concolor}, \emph{P. jeffreyi} and \emph{P. ponderosa} (SI Fig. 1). There are
also differences in drying rates across the eight species and estimated
mixed linear model slopes discriminated species into three groups from lowest
to most rapid: 1) \emph{C. decurrens}, 2) \emph{Abies spp} and \emph{S.
  giganteum}, and 3) \emph{Q. kelloggii} and \emph{Pinus spp} (SI Fig. 2).

@@

%% spread rate by moisture
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig4_spread_actualMC.pdf}
\caption{Flame spread rate by moisture content (model fit $F$= 61.15, df = 9,
  148, $P$ < 0.0001). Model ANOVA table in Table \ref{tab:spreadrate_anova} and
  coefficients in SI Table 4.}
  \label{fig:spread_moist}
\end{figure}

%% Table 3 
\begin{table}[h]
  \caption{Generalized linear mixed model results for two models predicting
    flame spread rate: the first model examined spread rate as a function of
    fuel moisture (mc) with taxonomic group as a fixed effect and species as a
    nested random effect. The second model predicted spread rate as a function
    of time since wetting (time) with taxonomic group as a fixed effect and
    species as a nested random effect. Approximate degrees of freedom (df),
    pseudo F statistics and p-values were calculated by the Kenward-Roger
    approximation \citep{Kenward_Roger-1997}. Estimated coefficients are in SI
    Tables 2 and 4}
  \label{tab:spreadrate_anova}

@@

%% fuel consumption by moisture:
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig5_consume_actualMC.pdf}
\caption{Percent fuel consumed by moisture content. Lines indicate best fit
  beta regression lines. Model deviance table in Table \ref{tab:consume_anova}
  and coefficients are in SI Table 3.}
  \label{fig:consume_moist}
\end{figure}

\begin{table}[h]
  \caption{Generalized linear mixed model results for two sets of models
    predicting fuel consumption. The first set modeled fuel consumption as a
    function of fuel moisture (mc) with taxonomic group as a fixed effect and
    species as a nested random effect. The second set modeled fuel consumption
    as a function of as a function of time since wetting (time) with taxonomic
    group as a fixed effect and species as a nested random effect. The analysis
    of deviance table below shows sequential chi-square tests on nested beta
    regression models: from the simplest null model within each group with only
    intercept and random effect, for a taxon effect only model, for a taxon and
    moisture content effect, and for the full model including the taxon by
    moisture or time interaction. Coefficient estimates are in SI Tables 3 and
    5.}
  \label{tab:consume_anova}
  \centering

@@

As a consequence of the relationships shown above, time since wetting had a
significant effect on spread rate ($P$ < 0.0001) and the slope of this effect
varied by taxonomic group (Fig. \ref{fig:spread_time}, Table
\ref{tab:spreadrate_anova}). The pines were able to ignite and burn much sooner
after wetting than did the other taxa, none of which had measurable spread
rates until burning was attempted 72 hours after wetting (note that trials that
failed to even ignite smoldering combustion are omitted completely from the
model and figure). At that point, the \emph{Abies}, \emph{Calocedrus}, and
\emph{Sequoiadendron} had dryer litter than did the \emph{Quercus}, yet the
spread rates were slow. \emph{Quercus} had the steepest response to time since
wetting (Fig. \ref{fig:spread_time}, SI Table 4)).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig6_spread_time.pdf}
\caption{Flame spread rate by time since wetting for litter of eight species.
  Lines indicate best fit linear model fits. Model ANOVA table in Table
  \ref{tab:spreadrate_anova} and coefficients in SI Table 4. Horizontal jitter
  (up to 2 hr) was added to the plot for visualization and to avoid
  overplotting. All collections were at the times labeled on the x-axis. A lack
  of data at earlier time points for a taxon indicates no ignitions occurred at
  those time points.}
\label{fig:spread_time}
\end{figure}

@@

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig7_consume_time.pdf}
\caption{Percent fuel consumed by time since wetting for litter of eight
  species. Curves shown indicate best fit beta regression lines. Model ANOVA
  table in Table \ref{tab:consume_anova} and coefficients in SI Table 4.
  Horizontal jitter (up to 2 hr) was added to the plot for visualization and to
  avoid overplotting. All collections were at the times labeled on the x-axis..
  A lack of data at earlier time points for a taxon indicates no ignitions
  occurred at those time points.}
  \label{fig:consume_time}
\end{figure}

@@

Mixture dry down rates and maximum water retention showed some relationship to
the dry down curves for individual constituent species (see SI Fig. 6). For
example, the mixture that does not contain \emph{Q. kelloggii} (AbCaPi in SI
Fig. 6) has the lowest initial retention. Initial water retention and rates
differed across mixtures (linear model based on log of moisture content, SI Table 8)

@@

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig9_mixture_obs_vs_pred_mc.pdf}
\caption{Observed vs predicted fuel moisture content for litter mixtures.
  Predictions are based on the mean moisture content at each time since wetting
  for the three constituent species in each mixture. Codes for species in
  mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_mc}
\end{figure}

@@

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig10_mixture_obs_vs_pred_spread.pdf}
\caption{Observed vs. predicted flame spread rate for litter mixtures. Filled
  circles show observed spread rate values. Open diamonds indicate predicted
  spread rates based on the average spread rate at each time since wetting for
  the three constituent species in each mixture. Codes for species in mixtures:
  Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_spread}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig11_mixture_obs_vs_pred_consume.pdf}
\caption{Observed vs predicted fuel consumption for litter mixtures. Filled
  circles show observed fuel consumption percentages. Open diamonds indicate
  predicted fuel consumption based on the average fuel consumption at each
  particular time since wetting for the three constituent species in each
  mixture. Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca =
  \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu =
  \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_consume}
\end{figure}

Flammability measures also demonstrated non-additivity in mixtures. Flame
spread rate was consistently higher than predicted (Fig
\ref{fig:mixture_obs_pred_spread}, Wald test of mixed linear model centered
intercept for observed-predicted values greater than zero, $P$ = 0.017) and the
positive non-additivity increased with time since wetting (slope greater than
zero, $P$ = 0.005). Fuel consumption, however, showed no overall bias towards
positive or negative non-additivity, but non-additivity increased with time
since wetting (Fig. \ref{fig:mixture_obs_pred_consume}, slope not equal to
zero, $P$ < 0.0001).

>>>>

Litter packing, measured as litter-bed density or packing ratio, controls the
flammability of dry litter fuels in seasonally dry climates because higher
litter density reduces flammability as a consequence of oxygen limitation. Leaf
litter density is driven by leaf size and shape
\citep{Fonda+Belanger+etal-1998, Scarff+Westoby-2006, Kane+Varner+etal-2008}.
These litter particle effects on litter bed aeration, in turn, have effects on
litter bed flammability \citep{Ganteaume+Marielle+etal-2011,
  Schwilk+Caprio-2011, Kreye+Varner+etal-2013}. These species-specific effects
are well established. However, how moisture dynamics modulate species specific
flammability is poorly known and could be driven by by leaf traits, emergent
litter litter traits, or both. Species specific effects on litter flammability
are recorded for the forests of the southeastern US \citep{Nowacki+Abrams-2008}
where species-specific litter drying rates play a role in flammability
feedbacks \citep{Kreye+Varner+etal-2013}. Similarly, in mixed-conifer forest of
the western USA, where the fire regime is characterized by surface fires fueled
by plant litter, overstory tree composition drives variation in local fire
behavior \citep{Schwilk+Caprio-2011} as a result of differential litter
flammability across tree species and as a result of synergistic interactions in
multi-species litter mixtures \citep{Magalhaes+Schwilk-2012}. These
flammability feedbacks may help explain historical fire frequency patterns, but
current estimates of litter flammability in mixed conifer forests are based
largely on burning dry litter.

@@

Field sites for this study were located in Sequoia and Kings Canyon National
Parks, California, USA (36 36N, 118 42 W) between 1600 and 2400 m elevation in
mixed-conifer forest. We chose eight dominant tree species that are
representative of this type of forest: \emph{Pinus jeffreyi} Grev. \& Balf.,
\emph{Pinus lambertiana} Dougl., \emph{Pinus ponderosa} Dougl. ex Laws.,
\emph{Abies concolor} (Gord. \& Glend.) Lindl. ex Hildebr., \emph{Abies
  magnifica} A. Murr., \emph{Calocedrus decurrens} (Torr.) Florin,
\emph{Quercus kelloggii} Newb., and \emph{Sequoiadendron giganteum} (Lindl.) J.
Buchholz. In this study, we consider litter the top layer of leaves and small
twigs less than 0.625 cm diameter (= 1-hour fuel) that have fallen in the
previous year (mostly undecomposed). Leaf litter was collected in the summer
(mid-June to mid-July) in 2012 from 21 separate sites across the study area. We
collected from a minimum of four populations per species with each population
9--32\,km distant from its neighbors. An exception to this was \emph{A.
  magnifica} which is restricted to the higher elevations and could only be
collected from one area due to designated wilderness boundaries and logistical
constraints. Within that area, \emph{A. magnifica} litter was collected from
over 10 different individuals. For the remaining species, the collection
involved 2--4 individual trees at each site which were separated from one
another by at least 10\,m. The litter was collected 2\,m away from the trunk to
obtain a more uniform sample of leaf and fine branchlet litter, because bark
and heavier branches tend to fall closer to the tree trunk. All litter was air
dried to $< 5$\% fuel moisture by dry weight for weighing and dividing into
moisture trials. The moisture manipulations and subsequent burning trials were
conducted in April--August 2015.

@@

Where $m$ is moisture content at time $t$, and $\lambda$ is the exponential
decay coefficient. To estimate these coefficients we fit a linear mixed-effects
model by first taking the natural logarithm of moisture content, with time
since wetting and species as fixed effects nested within replicate basket. We
extracted the coefficients of the fitted model to describe maximum water
retainability ($m_{max}$; g of water per g of dry mass) and desiccation rate
($\lambda$; hour$^{-1}$) for each species. All analyses were conducted in R 3.6.1
\citep{RCoreTeam-2019}. We fit models with R packages `lme4'
\citep{Bates_Machler_etal-2015} and `afex' \citep{Singmann_Bolker_etal-2017}.
Approximate degrees of freedom and p-values were calculated by the
Kenward-Roger approximation \citep{Kenward_Roger-1997} which is recommended by
Luke \citeyear{Luke-2017} as producing acceptable Type I error rates. We tested if
one leaf trait (specific leaf area) and one litter trait (litter bulk density)
influenced moisture dynamics by fitting linear models with species means as
observations. For each response variable we fit a single linear model with two
predictors and an interaction term.

Our past work \citep{Magalhaes+Schwilk-2012} demonstrated that dry litter
flammability across these eight species was driven by leaf traits and litter
packing. These previously reported flammability trials demonstrated that
species within a genus tended to behave similarly and the two scale-leaved
genera in the Cupressaceae, \emph{Calocedrus} and \emph{Sequoiadendron}, also
behaved similarly to one another. Therefore, to preserve power we used taxonomic
groups (\emph{Abies}, Cupressaceae, \emph{Quercus}, and \emph{Pinus}) as our
flammability functional groups and treated these as fixed effects in our
models. The effect of fuel moisture and time since wetting on flammability
parameters were modeled using linear models with fuel moisture or time since
wetting and taxon as fixed effects and average vapor pressure deficit in the
burning room during that trial as a nuisance covariate. We used linear mixed
models to predict spread rate, but fuel consumption tends to exhibit threshold
effects with few intermediate values. Therefore, we used mixed generalized
linear models with a beta regression link with the `glmmADMB' package
\citep{Skaug_Fournier_etal-2016} to model percent fuel consumed as a function of
moisture content and of time since wetting.

We investigated possible non-additive effects on moisture and on flammability
by testing if moisture content and flammability of mixtures were predicted by
the average of the constituent species. To choose mixtures, we first grouped
the species according to their maximum water retention and drying rate
according to estimated marginal means by species using the R package `emmeans'
\citep{Lenth-2019}. This resulted in four groups of species and we then
selected one representative species from each litter type group and produced
all four possible three-species litter mixtures using those four species. For
each mixture and time since wetting, we predicted moisture content based on the
mean content of the three constituent species at that time since wetting. We
predicted flammability parameters based on the average flammability measures
from individual species trials at each time since wetting and then calculated a
predicted spread rate and fuel consumption as the average of these three values
for the mixture. The expectation from the null model was that the difference
between observed and predicted values of flammability will be zero. We tested
if there were significant departures from the null with a mixed effects linear
model. For flammability parameters, we fit linear mixed effect models with the
observed-predicted flammability value as the dependent variable and used Wald
tests to test if the centered intercept was different than zero (a measure of
overall non-additivity) and if the slope with time since wetting was different
from zero (indicating non-additivity changing with drying).

@@

There were marked differences in the extent of moisture species retain
initially following wetting and draining by gravity (all $P < 0.0001$, Fig.
\ref{fig:drydown} and Table \ref{tab:drydown}). \emph{Q. kelloggii}, the only
broadleaf, is the species that retained the most moisture initially.
\emph{Quercus}, like the pines, had a rapid drying rate (Fig. \ref{fig:drydown}
and SI Fig. 1). Calculating estimated marginal means following the linear
mixed-effects model discriminated the species into three different maximum
water retention groups: 1) \emph{Q. kelloggii}; 2) \emph{P. lambertiana} and
\emph{C. decurrens}; and 3) \emph{A. magnifica}, \emph{S. giganteum}, \emph{A.
  concolor}, \emph{P. jeffreyi} and \emph{P. ponderosa} (SI Fig. 1). There are
also differences in drying rates across the eight species and estimated
mixed linear model slopes discriminated species into three groups from lowest
to most rapid: 1) \emph{C. decurrens}, 2) \emph{Abies spp} and \emph{S.
  giganteum}, and 3) \emph{Q. kelloggii} and \emph{Pinus spp} (SI Fig. 2).

@@

%% spread rate by moisture
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig2_spread_actualMC.pdf}
\caption{Flame spread rate by moisture content (model fit $F$= 61.15, df = 9,
  148, $P$ < 0.0001). Model ANOVA table in Table \ref{tab:spreadrate_anova} and
  coefficients in SI Table 4.}
  \label{fig:spread_moist}
\end{figure}

%% Table 3 
\begin{table}[h]
  \caption{Generalized linear mixed model results for two models predicting
    flame spread rate: the first model examined spread rate as a function of
    fuel moisture (mc) with taxonomic group as a fixed effect and species as a
    nested random effect. The second model predicted spread rate as a function
    of time since wetting (time) with taxonomic group as a fixed effect and
    species as a nested random effect. Approximate degrees of freedom (df),
    pseudo F statistics and p-values were calculated by the Kenward-Roger
    approximation \citep{Kenward_Roger-1997}. Estimated coefficients are in SI
    Tables 4 and 6}
  \label{tab:spreadrate_anova}

@@

%% fuel consumption by moisture:
\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig3_consume_actualMC.pdf}
\caption{Percent fuel consumed by moisture content. Lines indicate best fit
  beta regression lines. Model deviance table in Table \ref{tab:consume_anova}
  and coefficients are in SI Table 5.}
  \label{fig:consume_moist}
\end{figure}

\begin{table}[h]
  \caption{Generalized linear mixed model results for two sets of models
    predicting fuel consumption. The first set modeled fuel consumption as a
    function of fuel moisture (mc) with taxonomic group as a fixed effect and
    species as a nested random effect. The second set modeled fuel consumption
    as a function of as a function of time since wetting (time) with taxonomic
    group as a fixed effect and species as a nested random effect. The analysis
    of deviance table below shows sequential chi-square tests on nested beta
    regression models: from the simplest null model within each group with only
    intercept and random effect, for a taxon effect only model, for a taxon and
    moisture content effect, and for the full model including the taxon by
    moisture or time interaction. Coefficient estimates are in SI Tables 5 and
    7.}
  \label{tab:consume_anova}
  \centering

@@

As a consequence of the relationships shown above, time since wetting had a
significant effect on spread rate ($P < 0.0001$) and the slope of this effect
varied by taxonomic group (Fig. \ref{fig:spread_time}, Table
\ref{tab:spreadrate_anova}). The pines were able to ignite and burn much sooner
after wetting than did the other taxa, none of which had measurable spread
rates until burning was attempted 72 hours after wetting (note that trials that
failed to even ignite smoldering combustion are omitted completely from the
model and figure). At that point, the \emph{Abies}, \emph{Calocedrus}, and
\emph{Sequoiadendron} had dryer litter than did the \emph{Quercus}, yet the
spread rates were slow. \emph{Quercus} had the steepest response to time since
wetting (Fig. \ref{fig:spread_time}, SI Table 6)).

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig4_spread_time.pdf}
\caption{Flame spread rate by time since wetting for litter of eight species.
  Lines indicate best fit linear model fits. Model ANOVA table in Table
  \ref{tab:spreadrate_anova} and coefficients in SI Table 6. Horizontal jitter
  (up to 2 hr) was added to the plot for visualization and to avoid
  overplotting. All collections were at the times labeled on the x-axis. A lack
  of data at earlier time points for a taxon indicates no ignitions occurred at
  those time points.}
\label{fig:spread_time}
\end{figure}

@@

\begin{figure}[h]
  \centering
\includegraphics[width=8cm]{figs_tables/fig5_consume_time.pdf}
\caption{Percent fuel consumed by time since wetting for litter of eight
  species. Curves shown indicate best fit beta regression lines. Model ANOVA
  table in Table \ref{tab:consume_anova} and coefficients in SI Table 7.
  Horizontal jitter (up to 2 hr) was added to the plot for visualization and to
  avoid overplotting. All collections were at the times labeled on the x-axis..
  A lack of data at earlier time points for a taxon indicates no ignitions
  occurred at those time points.}
  \label{fig:consume_time}
\end{figure}

@@

Mixture dry down rates and maximum water retention showed some relationship to
the dry down curves for individual constituent species (see SI Fig. 6). For
example, the mixture that does not contain \emph{Q. kelloggii} (AbCaPi in SI
Fig. 6) has the lowest initial retention. Initial water retention and rates
differed across mixtures (linear model based on log of moisture content, SI Tables 8 and 9).

@@

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig6_mixture_obs_vs_pred_mc.pdf}
\caption{Observed vs predicted fuel moisture content for litter mixtures.
  Predictions are based on the mean moisture content at each time since wetting
  for the three constituent species in each mixture. Codes for species in
  mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_mc}
\end{figure}

@@

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig7_mixture_obs_vs_pred_spread.pdf}
\caption{Observed vs. predicted flame spread rate for litter mixtures. Filled
  circles show observed spread rate values. Open diamonds indicate predicted
  spread rates based on the average spread rate at each time since wetting for
  the three constituent species in each mixture. Codes for species in mixtures:
  Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi =
  \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_spread}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=16cm]{figs_tables/fig8_mixture_obs_vs_pred_consume.pdf}
\caption{Observed vs predicted fuel consumption for litter mixtures. Filled
  circles show observed fuel consumption percentages. Open diamonds indicate
  predicted fuel consumption based on the average fuel consumption at each
  particular time since wetting for the three constituent species in each
  mixture. Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca =
  \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu =
  \emph{Quercus kellogii}.}
  \label{fig:mixture_obs_pred_consume}
\end{figure}

Flammability measures also demonstrated non-additivity in mixtures. Flame
spread rate was consistently higher than predicted (Fig
\ref{fig:mixture_obs_pred_spread}, Wald test of mixed linear model centered
intercept for observed-predicted values greater than zero, $P$ = 0.017) and the
positive non-additivity increased with time since wetting (slope greater than
zero, $P$ = 0.005). Fuel consumption, however, showed no overall bias towards
positive or negative non-additivity, but non-additivity increased with time
since wetting (Fig. \ref{fig:mixture_obs_pred_consume}, slope not equal to
zero, $P < 0.0001$).

### 2019-08-06 ms/SI.tex
Correct SI fig and table numbers, add SI figures

\begin{table}[H]
  \caption{Mixed model coefficients for fuel consumption as a function of taxon and time since wetting. Table shows mixed linear model results.} 
  \label{tabS6}
\centering
\input{figs_tables/SI_tab7_consume_time_coef.ltx}
\end{table}

% table 7
\begin{table}[H]
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting for four different litter mixture types. Approximate
    degrees of freedom (df), pseudo F statistics and p-values were calculated
    by the Kenward-Roger approximation}.
  \label{tab:mixtures_drydown}
\centering

@@

@@

\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig4_mixture_drydown-curves.pdf}
\caption{Fuel moisture by time since saturation for four litter mixtures.  Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}. Moistures were measured every 24 hours but data are shown with slight horizontal jitter to reduce point overlap. Mixed linear model results shown in SI Table 7.}
  \label{SI_fig4}
\end{figure}

>>>>

\begin{table}[H]
  \caption{Mixed model coefficients for fuel consumption as a function of taxon and time since wetting. Table shows mixed linear model results.} 
  \label{tabS7}
\centering
\input{figs_tables/SI_tab7_consume_time_coef.ltx}
\end{table}

% table 8
\begin{table}[H]
  \caption{Linear mixed model results for moisture content as a function of
    time since wetting for four different litter mixture types. Approximate
    degrees of freedom (df), pseudo F statistics and p-values were calculated
    by the Kenward-Roger approximation}.
  \label{tab:mixtures_drydown}
\centering

@@

\begin{table}[H]
  \caption{Linear mixed model coefficients for moisture content as a function of
    time since wetting for four different litter mixture types.}.
  \label{tab:mixtures_drydown}
\centering
\begin{tabular}{rrrrrr}
  \toprule
 & Estimate & Std. Error & df & t value & Pr($>$$|$t$|$) \\ 
  \midrule
(Intercept) & 4.59 & 0.06 & 71.04 & 77.18 & 0.00 \\ 
  hour & -0.01 & 0.00 & 95.45 & -13.96 & 0.00 \\ 
  spcodeAbCaQu & 0.26 & 0.08 & 95.35 & 3.28 & 0.00 \\ 
  spcodeAbPiQu & 0.54 & 0.08 & 96.53 & 7.04 & 0.00 \\ 
  spcodeCaPiQu & 0.57 & 0.08 & 95.35 & 7.25 & 0.00 \\ 
  hour:spcodeAbCaQu & 0.00 & 0.00 & 95.40 & 3.67 & 0.00 \\ 
  hour:spcodeAbPiQu & -0.00 & 0.00 & 95.78 & -0.70 & 0.49 \\ 
  hour:spcodeCaPiQu & 0.00 & 0.00 & 95.40 & 2.64 & 0.01 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[H]
  \centering
\includegraphics[width=16cm]{figs_tables/SI_fig4_SLA_maxMC.pdf}
\caption{Maximum water retention by specific leaf area (SLA) across eight species.}
  \label{fig:maxmc-di}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=16cm]{figs_tables/SI_fig5_di_bd.pdf}
\caption{Linear model predicting dessication rate as a function of specific leaf area and litter bulk density across eight species.}.
  \label{fig:bd-di}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig6_mixture_drydown-curves.pdf}
\caption{Fuel moisture by time since saturation for four litter mixtures.  Codes for species in mixtures: Ab = \emph{Abies concolor}, Ca = \emph{Calocedrus decurrens}, Pi = \emph{Pinus jeffreyi} and Qu = \emph{Quercus kellogii}. Moistures were measured every 24 hours but data are shown with slight horizontal jitter to reduce point overlap. Mixed linear model results shown in SI Table 7.}
  \label{SI_fig6}
\end{figure}
